Lucky Thirteen - A Cryptographic Analyses

### 2013-07-14 paper.tex
Section V alpha

%SECTION I
\section{Introduction}
\label{sec:intro}
\textcolor{red}{TODO: fix introduction, make more clear what TLS is used for and what makes this attack so dangerous}

This paper explores the February 2013 theoretical attack on TLS (and DTLS) called \textit{Lucky Thirteen}~\cite{alfardan2013lucky}. This theoretic attack describes why the bandits used to fix TLS 1.1 and 1.2 after the attack of~\citeauthor{canvel2003password}~\cite{canvel2003password} do not provide the necessary security to the most used internet protocol.This paper is scoped to the implications of this attack on TLS Security, and does not provide information on how to apply a similar attack on DTLS. The reader is referred to~\cite{alfardan2013lucky} for details on DTLS, although mentionable is the fact that the theoretical attack on TLS, is practical for DTLS due to its nature not to terminate sessions on bad MAC or bad padding errors.

@@

% SECTION II-A
\subsection{Cryptographic format}
\label{sec:crypto:format}
The cryptographic format used in this paper is presented in Table~\ref{sec:crypto:format:table}.
\begin{table}[h]
    \begin{tabular}{l|l}
    Format & Meaning \\ \hline
    A $||$ B    & Concatenation of A and B  \\
    A $\oplus$ B  & The bitwise XOR of A and B     \\
    H(A)    & Hash-function applied to A    \\
    $E_k$(A)    & Encryption with key K applied to A    \\
    $D_k$(A)    & Decryption with key K applied to A    \\
    \end{tabular}
    \caption{Cryptographic format used in this paper}
    \label{sec:crypto:format:table}
\end{table}

% SECTION II-B
\subsection{MAC - Message Authentication Code}
\label{sec:crypto:hmac}
To ensure message integrity, RFC5246~\cite{ietf2008transport} prescribes the use of a keyed Message Authentication Code (MAC). The MAC cipher suites described in the RFC are based on hashing functions, HMACs, and are described in~\cite{krawczyk1997rfc}. Though, the RFC does not prohibit the use of other cipher suites, the implementations found by~\citeauthor{alfardan2013lucky} in~\cite{alfardan2013lucky} all restrict the use to HMAC-MD5, HMAC-SHA-1 or HMAC-SHA-256~\footnote{HMAC-MD5 has 16 bytes of MAC.\\HMAC-SHA-1 has 20 bytes of MAC.\\HMAC-SHA-256 has 32 bytes of MAC.} To compute the HMAC for a message M and key K, the specified HMAC algorithms use the following formula. 
\[ \text{HMAC}(K,M) = H((K \oplus opad)||H(K \oplus ipad)||M)). \]
With \textit{opad}, \textit{ipad} to be specific $64$-byte values and key $K$ zero-padded to $64$ bytes. "Each hash function uses a \textit{Merkle-Damg\aa rd strengthening}. Here, an 8-byte length field followed by padding of a specified byte format are appended to the message M to be hashed. The padding is at least $1$ byte in length and aligns the data on a $64$-byte boundary. The relevant hash functions also have an iterated structure, processing messages in chunks of $64$ bytes ($512$ bits) using a compression function, with the output of each compression step being chained into the next step. The compression function in turn involves a complex round structure, with many basic arithmetic operations on data being involved in each round."~\cite{alfardan2013lucky}.

\textcolor{red}{TODO: more elaborate; maybe a figure or a formula with regards to the above Merkle-Damg\aa rd strengthening.}

This means that, depending on the size of message M, the number of runs of the hash function H can be computed. Because of the Merkle-Damg\aa rd strengthening, a message M is appended by the $8$-byte length field, and padded to be a multiple of $64$ bytes. This means that the first of the Merkle-Damg\aa rd rounds consists of $55$ bytes of message M, followed by $64$ bytes for every following $64$ bytes of message M. The inner hash function depends on the $\text{message length of} M + 1$ constant hash function round, the outer hash function is a constant of $2$. This means that the following formula represents the number of hash function calls for L is the message length of M in bytes:
\[ \lceil \frac{L - 55}{64} \rceil + 4 \]

@@

% SECTION III-B
\subsection{Accepting a record}
\label{sec:tls:accepting}
The RFC~\cite{ietf2008transport} does not describe the steps that need to be taken before accepting the payload of a received TLS record, fortunately the authors of the Lucky Thirteen attack have proposed a (incomplete) list of actions to be undertaken.
\begin{itemize}
  \item Check: is the ciphertext size a multiple of the block size?
  \item Check: is the ciphertext large enough to contain at least a zero-length record, a MAC tag of the required size and at least one byte of padding?
  \item Decrypt the ciphertext to recover the plaintext blocks: $P_i = D_k(C_i) \oplus C_{i-1}$.
  \item Is the padding of the appropriate format?
  \item Is the MAC correct (using the correct SQN number and HDR field)?
\end{itemize}
Since the attack by~\citeauthor{vaudenay2002security}~\cite{vaudenay2002security} on TLS 1.0, exploiting differences when accepting records, it is advised by TLS 1.1 and 1.2~\cite{ietf2008transport} to avoid timing differences by assuming zero length padding if the padding format sanity check does not work out. In this case, the MAC is still calculated, but the record is nonetheless discarded. More on this in Section~\ref{sec:paddingoracle} and Section~\ref{sec:lucky}.

@@

% SECTION IV-A
\subsection{~\citeauthor{vaudenay2002security} attack on TLS 1.0}
\label{sec:paddingoracle:padding}
For this Section, w.l.o.g., it will be assumed that this application data protocol run uses the HMAC\_SHA1 with AES\_CBC mode~\footnote{This makes for t = 20 bytes MAC size. \\ b = 16 bytes block size}.\footnote{The format in this Section mostly depends upon the same format as used in~\cite{alfardan2013lucky}, to keep consistency throughout the literature. Furthermore, this subsection is heavily inspired by ~\cite{alfardan2013lucky,vaudenay2002security}.}

@@

@@

The formal algorithms are as follows.
\begin{verbatim}
Decrypt last byte
1. Set zb = 0x00, delta = (16 times) zb.
2. Construct C_att = HDR||IV||C_1||C_2||
                C' xor delta ||C*.
3. Send C_att to the Oracle.
4. If 'decryption_failed' is thrown then
6. -Set delta = delta + 0x01.
6. -Go to step 2.
7. Else
8. -Set delta = delta xor ((14 times) zb
                ||0xFF||0xFF).
9. -Construct C_att = HDR||IV||C_1||C_2||
                C' xor delta||C*.
10. -Send C_att to the Oracle.
11. -If 'decryption_failed' is thrown then
12. --Go to step 6.
13. -Else
14. --output delta.
15. -Endif.
16. Endif.
17. Set delta = delta xor ((14 times) zb
                 || 0xFF || zb.
18. Go to step 2.
\end{verbatim}
The output is the $\Delta$ for which $P_4$ ends in $0\text{x}00$. The last byte of $P^{*}$ is equal to the last byte of $P_4 \oplus \Delta$. Since the last byte of $P_4$ is $0\text{x}00$, the last byte of $P^{*}$ is equal to the last byte of $\Delta$.

To get the plaintext byte i from the end (starting with the last byte i = 0), consider $rm^{'}$ to be the $\Delta$ for which the padding is correct up to a padding length of i+1 (thus, the content of the padding being i). Set $rm$ to be $rm^{'}$ with the last i bytes all added up with 1.
\begin{verbatim}
Decrypt byte i (from the end)
i bytes from the end known
(because counting from 0)
1. Set zb = 0x00, j = 0.
2. Set delta = (16 times) zb xor rm.
3. Set j = j + 1.
4. Set delta = delta xor (15-i times) zb 
          || zb + j || (i times) 0xFF.
5. Construct C_att = HDR||IV||C_1||C_2||
          C' xor delta||C*.
6. Send C_att to the Oracle.
7. If 'decryption_failed' is thrown then
8. -Go to step 3.
9. Else
10. -Set delta = delta xor 
         (15-i-1 times) zb || 
         (i+2 times) 0xFF
11. -Construct C_att = HDR||IV||C_1||C_2||
          C' xor delta||C*.
12. -Sent C_att to the Oracle.
13. -If 'decryption_failed' is thrown then
14. --Go to step 4.
15. -Else
16. --output delta.
17. -Endif.
18. Endif.
19. Set delta = delta xor ((15-i-1 
                 times) zb || 0xFF || 
                 (i + 1 times) zb.
20. Set j = 0.
21. Go to step 3.
\end{verbatim}
The output is the $\Delta$ for which $P_4$ ends in i+1 times (counting from 1) $0\text{x}i$. The last i+1 bytes of $P^{*}$ are equal to the last i+1 bytes of $P_4 \oplus \Delta$.

Using this last algorithm, one can completely decrypt ciphertext $C^{*}$ into $P^{*}$, with an average of $\frac{2^8}{2} = 2^7$ tries per byte, having $16$ ($2^4$) bytes in AES\_CBC mode, the complete average of recovering block $C^{*}$ is $2^7 * 2^4 = 2^{11}$.

@@

The Lucky Thirteen attack will be described in three parts, first the bandit of applying zero length padding is discussed, after which the complexity of the attack is explained, and finally the practicality of this attack is discussed.

@@

% SECTION V-B
\subsection{Complexity of Lucky Thirteen}
\label{sec:lucky:complexity}
As shown in Section~\ref{sec:lucky:bandit}, the case in which a valid padding pattern can be distinguished, needs at least a valid padding pattern of length two. This leads to complexity of (in worst case), trying all byte variations for the last two bytes of $\Delta$, leading to a worst case complexity of $2^8*2^8 = 2^{16}$, in the average case $\frac{2^{16}}{2} = 2^{15}$. With knowledge about the plaintext $P^{*}$'s last two bytes, the third-to-last byte can be discovered by setting the last two plaintext bytes of $P_4$ to $0\text{x}02||0\text{x}02$ and altering the third-to-last byte. In worst case, this take $2^8$ trials with an average of $\frac{2^8}{2} = 2^7$ for every byte of plaintext from right to left. To recover an entire plaintext block $P^{*}$, one needs to first get the last two bytes with a complexity of $2^{15}$ and then find the other $14$ bytes with complexity $2^7$, giving a total of $2^{15} * 14 * 2^7 \approx 2^{26}$ trials.

% SECTION V-C
\subsection{Practicality of Lucky Thirteen}
\label{sec:lucky:practicality}
\textcolor{red}{First explain the results from Lucky Thirteen, than discuss how to build the attack using multi-channel}

Lucky Thirteen~\cite{alfardan2013lucky}

Multi-channel attack~\cite{canvel2003password}

BEAST~\cite{duong2011here}

% SECTION VI
\section{The Future of TLS}
\label{sec:future}

>>>>

%SECTION I
\section{Introduction}
\label{sec:intro}
TLS, Transport Layer Security, is the most used cryptographic protocol to encrypt end-to-end connections on Transport Layer level. It is used when browsing via HTTPS (HTTP Secure), for VoIP, e-mail, and lots of other applications. The core of this program is the end-to-end encryption (using symmetric cryptography) that it allows after an authentication (using asymmetric cryptography). TLS can be used to prevent eavesdropping and tampering in communication in a client-server application. One detailed example of usage is when using the stateless HTTP protocol, which uses \textit{cookies} to fake state. The connection in which cookies are sent from a client to the server is protected using TLS to avoid tampering and eavesdropping on the cookie. If an attacker would be able to steal a(nother) clients cookie, he could fake being that user to the server, potentially being able to login to banking accounts, e-mail applications and many other applications.

This paper explores the February 2013 theoretical attack on TLS (and DTLS) called \textit{Lucky Thirteen}~\cite{alfardan2013lucky}. This theoretic attack describes why the bandits used to fix TLS 1.1 and 1.2 after the attack of~\citeauthor{canvel2003password}~\cite{canvel2003password} do not provide the necessary security to the most used internet security protocol.This paper is scoped to the implications of this attack on TLS Security, and does not provide information on how to apply a similar attack on DTLS. The reader is referred to~\cite{alfardan2013lucky} for details on DTLS, although mentionable is the fact that the theoretical attack on TLS, is practical for DTLS due to its nature not to terminate sessions on bad MAC or bad padding errors.

@@

% SECTION II-A
\subsection{Cryptographic format}
\label{sec:crypto:format}
The cryptographic format used in this paper is presented in Table~\ref{sec:crypto:format:table}.
\begin{table}[h]
    \begin{tabular}{l|l}
    Format & Meaning \\ \hline
    A $||$ B    & Concatenation of A and B  \\
    A $\oplus$ B  & The bitwise XOR of A and B     \\
    H(A)    & Hash-function applied to A    \\
    $E_k$(A)    & Encryption with key K applied to A    \\
    $D_k$(A)    & Decryption with key K applied to A    \\
    $P_{14}$    & $14$th byte of $P$ \\
    \end{tabular}
    \caption{Cryptographic format used in this paper}
    \label{sec:crypto:format:table}
\end{table}

% SECTION II-B
\subsection{MAC - Message Authentication Code}
\label{sec:crypto:hmac}
To ensure message integrity, RFC5246~\cite{ietf2008transport} prescribes the use of a keyed Message Authentication Code (MAC). The MAC cipher suites described in the RFC are based on hashing functions, HMACs, and are described in~\cite{krawczyk1997rfc}. Though, the RFC does not prohibit the use of other cipher suites, the implementations found by~\citeauthor{alfardan2013lucky} in~\cite{alfardan2013lucky} all restrict the use to HMAC-MD5, HMAC-SHA-1 or HMAC-SHA-256~\footnote{HMAC-MD5 produces 16 bytes of MAC.\\HMAC-SHA-1 produces 20 bytes of MAC.\\HMAC-SHA-256 produces 32 bytes of MAC.} To compute the HMAC for a message M and key K, the specified HMAC algorithms use the following formula. 
\[ \text{HMAC}(K,M) = H((K \oplus opad)||H(K \oplus ipad)||M)). \]
With \textit{opad}, \textit{ipad} to be specific $64$-byte values and key $K$ zero-padded to $64$ bytes. "Each hash function uses a \textit{Merkle-Damg\aa rd strengthening}. Here, an 8-byte length field followed by padding of a specified byte format are appended to the message M to be hashed. The padding is at least $1$ byte in length and aligns the data on a $64$-byte boundary. The relevant hash functions also have an iterated structure, processing messages in chunks of $64$ bytes ($512$ bits) using a compression function, with the output of each compression step being chained into the next step. The compression function in turn involves a complex round structure, with many basic arithmetic operations on data being involved in each round."~\cite{alfardan2013lucky}.\footnote{The exact working of this Merkle-Damg\aa rd strengthening is not important for the Lucky Thirteen and therefore falls out of the scope of this paper.}
This means that, depending on the size of message M, the number of runs of the hash function H can be computed. Because of the Merkle-Damg\aa rd strengthening, a message M is appended by the $8$-byte length field, and padded to be a multiple of $64$ bytes. This means that the first of the Merkle-Damg\aa rd rounds consists of $55$ bytes of message M, followed by $64$ bytes for every following $64$ bytes of message M. The inner hash function depends on the $\text{message length of} M + 1$ constant hash function round, the outer hash function is a constant of $2$. This means that the following formula represents the number of hash function calls for L is the message length of M in bytes:
\[ \lceil \frac{L - 55}{64} \rceil + 4 \]

@@

% SECTION III-B
\subsection{Accepting a record}
\label{sec:tls:accepting}
The RFC~\cite{ietf2008transport} does not describe the steps that need to be taken before accepting the payload of a received TLS record, fortunately the authors of the Lucky Thirteen attack have proposed a (incomplete) list of actions to be undertaken.
\begin{itemize}
  \item Check: is the ciphertext size a multiple of the block size?
  \item Check: is the ciphertext large enough to contain at least a zero-length record, a MAC tag of the required size and at least one byte of padding?
  \item Decrypt the ciphertext to recover the plaintext blocks: $P_i = D_k(C_i) \oplus C_{i-1}$.
  \item Check: is the padding of the appropriate format?
  \item Check: is the MAC correct (using the correct SQN number and HDR field)?
\end{itemize}
Since the attack by~\citeauthor{vaudenay2002security}~\cite{vaudenay2002security} on TLS 1.0, exploiting differences when accepting records, it is advised by TLS 1.1 and 1.2~\cite{ietf2008transport} to avoid timing differences by assuming zero length padding if the padding format sanity check does not work out. In this case, the MAC is still calculated, but the record is nonetheless discarded. More on this in Section~\ref{sec:paddingoracle} and Section~\ref{sec:lucky}.

@@

% SECTION IV-A
\subsection{~\citeauthor{vaudenay2002security} attack on TLS 1.0}
\label{sec:paddingoracle:padding}
For this Section, w.l.o.g., it will be assumed that this application data protocol run uses the HMAC\_SHA1 with AES\_CBC mode~\footnote{This makes for t = 20 bytes MAC size and b = 16 bytes block size}.\footnote{The format in this Section mostly depends upon the same format as used in~\cite{alfardan2013lucky}, to keep consistency throughout the literature. Furthermore, this subsection is heavily inspired by ~\cite{alfardan2013lucky,vaudenay2002security}.}

@@

Figure~\ref{sec:paddingoracle:padding:lastbyte} shows the algorithm used to decrypt the last byte.

@@

\begin{figure}
\begin{verbatim}
Decrypt the last byte
1. Set zb = 0x00, delta = (16 times) zb.
2. Construct C_att = HDR||IV||C_1||C_2||
                C' xor delta ||C*.
3. Send C_att to the Oracle.
4. If 'decryption_failed' is thrown then
6. -Set delta = delta + 0x01.
6. -Go to step 2.
7. Else
8. -Set delta = delta xor ((14 times) zb
                ||0xFF||0xFF).
9. -Construct C_att = HDR||IV||C_1||C_2||
                C' xor delta||C*.
10. -Send C_att to the Oracle.
11. -If 'decryption_failed' is thrown then
12. --Go to step 6.
13. -Else
14. --output delta.
15. -Endif.
16. Endif.
17. Set delta = delta xor ((14 times) zb
                 || 0xFF || zb.
18. Go to step 2.
\end{verbatim}
\caption{Algorithm to decrypt the last byte}
\label{sec:paddingoracle:padding:lastbyte}
\end{figure}
In Figure~\ref{sec:paddingoracle:padding:lastbyte}, the output is the $\Delta$ for which $P_4$ ends in $0\text{x}00$. The last byte of $P^{*}$ is equal to the last byte of $P_4 \oplus \Delta$. Since the last byte of $P_4$ is $0\text{x}00$, the last byte of $P^{*}$ is equal to the last byte of $\Delta$.

To get the plaintext byte i from the end (starting with the last byte i = 0), consider $rm^{'}$ to be the $\Delta$ for which the padding is correct up to a padding length of i+1 (thus, the content of the padding being i). Set $rm$ to be $rm^{'}$ with the last i bytes all added up with 1.

Figure~\ref{sec:paddingoracle:padding:bytei} shows the algorithm to decrypte byte i from the end.

\begin{figure}
\begin{verbatim}
Decrypt byte i (from the end)
i bytes from the end known
(because counting from 0)
1. Set zb = 0x00, j = 0.
2. Set delta = (16 times) zb xor rm.
3. Set j = j + 1.
4. Set delta = delta xor (15-i times) zb 
          || zb + j || (i times) 0xFF.
5. Construct C_att = HDR||IV||C_1||C_2||
          C' xor delta||C*.
6. Send C_att to the Oracle.
7. If 'decryption_failed' is thrown then
8. -Go to step 3.
9. Else
10. -Set delta = delta xor 
         (15-i-1 times) zb || 
         (i+2 times) 0xFF
11. -Construct C_att = HDR||IV||C_1||C_2||
          C' xor delta||C*.
12. -Sent C_att to the Oracle.
13. -If 'decryption_failed' is thrown then
14. --Go to step 4.
15. -Else
16. --output delta.
17. -Endif.
18. Endif.
19. Set delta = delta xor ((15-i-1 
                 times) zb || 0xFF || 
                 (i + 1 times) zb.
20. Set j = 0.
21. Go to step 3.
\end{verbatim}
\caption{Algorithm to decrypt byte i (from the end)}
\label{sec:paddingoracle:padding:bytei}
\end{figure}

In Figure~\ref{sec:paddingoracle:padding:bytei}, the output is the $\Delta$ for which $P_4$ ends in i+1 times (counting from 1) $0\text{x}i$. The last i+1 bytes of $P^{*}$ are equal to the last i+1 bytes of $P_4 \oplus \Delta$.

Using the algorithm in Figure~\ref{sec:paddingoracle:padding:bytei}, one can completely decrypt ciphertext $C^{*}$ into $P^{*}$, with an average of $\frac{2^8}{2} = 2^7$ tries per byte, having $16$ ($2^4$) bytes in AES\_CBC mode, the complete average of recovering block $C^{*}$ is $2^7 * 2^4 = 2^{11}$.

@@

The Lucky Thirteen attack will be described in three parts, first the bandit of applying zero length padding is discussed, after which the complexity of the attack is explained, thereafter the practicality of this attack is discussed, ending in a subsection which point to some possible improvements of the attack.

@@

% SECTION V-B
\subsection{Complexity of Lucky Thirteen}
\label{sec:lucky:complexity}
As shown in Section~\ref{sec:lucky:bandit}, the case in which a valid padding pattern can be distinguished, needs at least a valid padding pattern of length two. This leads to complexity of (in worst case), trying all byte variations for the last two bytes of $\Delta$, leading to a worst case complexity of $2^8*2^8 = 2^{16}$, in the average case $\frac{2^{16}}{2} = 2^{15}$. With knowledge about the plaintext $P^{*}$'s last two bytes, the third-to-last byte can be discovered by setting the last two plaintext bytes of $P_4$ to $0\text{x}02||0\text{x}02$ and altering the third-to-last byte. In worst case, this take $2^8$ trials with an average of $\frac{2^8}{2} = 2^7$ for every byte of plaintext from right to left. To recover an entire plaintext block $P^{*}$, one needs to first get the last two bytes with a complexity of $2^{15}$ and then find the other $14$ bytes with complexity $2^7$, giving a total of $2^{15} + 14 * 2^7 = 34560 \approx 2^{15}$ trials on average.

% SECTION V-C
\subsection{Practicality of Lucky Thirteen}
\label{sec:lucky:practicality}
The actual novelty of the Lucky Thirteen attack is in the fact that~\citeauthor{alfardan2013lucky} proved that (within a LAN), the time difference of one hash round can be approached after only a small amount of duplicate session runs. The exact details for their setup can be found in~\cite{alfardan2013lucky}, but it boils down to a regular client-server architecture with regular hardware in a LAN. Their statistical analyses for distinguishing whether or not the extra hash round was done showed that with a success probability of 1, only $2^7$ duplicate session runs. A complete set of results for this distinguishing attack can be found in Table~\ref{sec:lucky:practicality:distinguishtable}\footnote{OpenSSL was used for this setup}, L is the number of duplicate session runs.

\begin{table}[h]
    \begin{tabular}{|l|l|}
    \hline
    L & Success Probability \\ \hline \hline
    1 & 0.756  \\ \hline
    2 & 0.769  \\ \hline
    4 & 0.858  \\ \hline
    8 & 0.914  \\ \hline
    16 & 0.951  \\ \hline
    32 & 0.983  \\ \hline
    64 & 0.992  \\ \hline
    128 & 1 \\ \hline
    \end{tabular}
    \caption{Distinguishing attack success probabilities~\cite{alfardan2013lucky}}
    \label{sec:lucky:practicality:distinguishtable}
\end{table}

\citeauthor{alfardan2013lucky} furthermore did experiments for finding one of the last two bytes, if the other was already known. In this experiment, they knew $P_{14}$ (the penultimate byte) and set it to $0\text{x}01$, and tried to find $\Delta_{15}$ so that the plaintext byte $P_{15}$ would turn out to be $0\text{x}01$. Figure~\ref{fig:luckybyteserver} shows the hardware cycles that were calculated by the server, while Figure~\ref{fig:luckybyteattacker} shows the same calculation seen from the attackers point of view, including all (network) noise in the LAN / hardware. The same setup was used as described for the distinguishability attack above, with $L$ being $2^7$. As can be seen clearly, the attacker can easily distinguish the value for $\Delta_{15}$ which turns $P_{15}$ into $0\text{x}01$, namely $\Delta_{15} = 0\text{xFE}$.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth]{lucky_byte_server.png}
	\caption{OpenSSL TLS median network timings in terms of hardware cycles for $\Delta_{15}$ being $0\text{xFE}$, as calculated by the server.~\cite{alfardan2013lucky}}
	\label{fig:luckybyteserver}
\end{figure}

\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth]{lucky_byte_attacker.png}
	\caption{OpenSSL TLS median network timings in terms of hardware cycles for $\Delta_{15}$ being $0\text{xFE}$, as calculated by the attacker.~\cite{alfardan2013lucky}}
	\label{fig:luckybyteattacker}
\end{figure}

Considering that an attacker wants to decrypt a complete ciphertext block, Section~\ref{sec:lucky:complexity} already showed that, without connection failures, it would take approximately $2^{15}$ trails. Now, taken into account the number of connection necessary to resolve a complete byte with a probability of $1$, L = $2^7$, the attacker needs about $2^{22}$ TLS connections on average. \citeauthor{alfardan2013lucky} calculated that, with their setup, they need about $32$ hours because of the many TLS connection setups and teardowns.

% Section V - D
\subsection{Attack improvements}
\label{sec:lucky:improvements}
The authors of the Lucky Thirteen~\cite{alfardan2013lucky} attack propose some improvements to their attack, which they did not implement nor test. The improvements are met due to the Padding Oracle that was explained in Section~\ref{sec:paddingoracle:bomb} in which the attacker can query for the validity of a plaintext assumption. In the BEAST attack~\cite{duong2011here}, ~\citeauthor{duong2011here} describe that when querying for such plaintext assumptions, the implementation of the HTTP cookies come in very handy since they are encoded using base64 encoding. Base64 encoding produces bytes between [$1..2^6$], meaning that querying for cookie bytes needs only $\frac{L*2^{6}}{2} = 2^7 * 2^5 = 2^{12}$ TLS connections on average when the last two bytes are discovered. These two bytes take $\frac{L*2^6*2^6}{2} = 2^7*2^6*2^5 = 2^{18}$ TLS connections, summing up to $2^{12} + 2^{18} = 266240 \approx 2^{25}$ TLS connections.

Other improvements using the same technique can be used to order the set of plaintext assumptions according to statistical analysis of i.e. the language used in the plaintext blocks, or the format (i.e. XML / HTML). Consider a message of which it is assumed to contain English text, the set of bytes that that block could contain, is highly biased towards ASCII text representable bytes. These improvements can speed up the attack considerably.

% SECTION VI
\section{The Future of TLS}
\label{sec:future}
\textcolor{red}{TODO: Explain the fixes that could be introduced, also check for newer papers (that  cite~\citeauthor{alfardan2013lucky}).}
